---
title: "data preprocessing notes"
output: html_notebook
---
Objective: A curated list of handy `base R` functions for data preprocessing tasks. This list can be used for interview questions and answers.

```{r}
data("airquality")

# check data dimensions
dim(airquality)

# check missing values
colSums(is.na(airquality))
```
Let's look at the data structure
```{r}
str(airquality)
```

The variables `Ozone` and `Solar.R` contain missing values. Both of them are continuous in nature. So to impute the missing data, we can use the `median`.
```{r}
for(i in 1:ncol(airquality)){
  airquality[is.na(airquality[,i]), i] <- median(airquality[,i], na.rm = TRUE)
}
```
Check for missing values again.
```{r}
colSums(is.na(airquality))
```
**Divide the data into training and testing/validation set**
```{r}
## 75% will be training data, rest be test data
smp_size <- floor(0.75 * nrow(airquality))

## set the seed to make your partition reproducible
set.seed(2018)
train_ind <- sample(seq_len(nrow(airquality)), size = smp_size)

train_data <- airquality[train_ind, ]
test_data <- airquality[-train_ind, ]
dim(train_data)
dim(test_data)
```

** A classification task**: Let's apply a machine learning technique like regression. Note the base R function, `lm()` can both be used as a regression tool and feature selection tool.
```{r}
# assume we want to build model over day
linear_model<- lm(Day ~., data = train_data)
print(linear_model)
```
*** Linear Regression Diagnostics***
```{r}
summary(linear_model)
```
The `summary()` generates a lot of information. For now, our concern is `feature selection`. The number of stars besides a given variable indicate its importance. So we can see, the variable `Solar.R` is the only significant variable to accurately predict `Day`.
Let's try another variab;e like `Month` to predict.
```{r}
linear_model<- lm(Month~., data = airquality)
summary(linear_model)
```
